---
title: "GlmsRedefined"
output: html_document
date: "2024-06-28"
---

```{r setup, include=FALSE}
# MuMIn: For model selection based on Akaike Criterion (AIC)
if (!require("MuMIn")) install.packages("MuMIn")
# car for collinearity
if (!require("car")) install.packages("car")
# formatting nas
if (!require("dplyr")) install.packages("dplyr")
# colinearity with checkcolinearity
if (!require("performance")) install.packages("performance")
# summary with j.tools
if (!require("jtools")) install.packages("jtools")
# for post-hocs
if (!require("emmeans")) install.packages("emmeans")

# Package loading
library(MuMIn)
library(car)
library(dplyr)
library(performance) # checkcollinearity
library(jtools) #summ
library(emmeans) #post-hocs
```

## Loading data
```{r}
library(readxl)
Data1 <- read_excel("~/Desktop/Manuscrito/Data/Datafinal/nuevejul/databasesinhum.xlsx")
# NAs ya gesionados, filas eliminadas a mano
```

## PHYSICAL Aggression 
            and EAE
```{r}
# Converting Physical Aggression to z-scores
z.PhysicalAggression <- as.vector(scale(Data1$PhysicalAggression))

# Full model
fullPHY <- glm(z.PhysicalAggression ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullPHY) # [1] 122.0248
check_collinearity(fullPHY) # >10

# Best model
bestPHY <- dredge(fullPHY,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestPHY, delta <= 4)
model.avg(bestPHY, subset = cumsum(weight) <= .95)
summary(get.models(bestPHY, 1)[[1]]) # Al aplicar la función summary te da directacmente p-values para subcategorías de las variables. Revisando el artículo de Amapola, veo que al final sí lo reportó así. ver por ejemplo su Tabla 17.

# Simplified summary
library(jtools) 
summ(BestmodelPHY, vifs=TRUE) # Una función parecida al summary, que te da los mismos valores, pero con solo 2 decimales. Amapola copió sus tablas de resultados a partir de esto.

BestmodelPHY <- (get.models(bestPHY, 1)[[1]])
AICc(BestmodelPHY) # [1] 111.196 (Menor que el del full model, OK)

#VIFS:
check_collinearity(BestmodelPHY) # = 2.74

#POST-HOCS:
emmeans (BestmodelPHY, pairwise ~ Abuse)
#los p-values indican que hay diferencias significativas entre B1-B4 y B3-B4. /// Explicacion: Si la comparación "None" vs "High" muestra una diferencia significativa, implica que las personas en el nivel "High" de Abuse tienen valores de z.PhysicalAggression significativamente diferentes (mayores o menores) en comparación con las personas en el nivel "None".
emmeans (BestmodelPHY, pairwise ~ AgeArrival)
#los p-values indican que hay diferencias significativas entre A2-A4 y A3-A4
```
## SOCIAL Aggression 
            and EAE
```{r}
# Converting Social Aggression to z-scores
z.SocialAggression <- as.vector(scale(Data1$SocialAggression))

# Full model
fullSOC <- glm(z.SocialAggression ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullSOC) # [1] 136.053

# Best model
bestSOC <- dredge(fullSOC,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestSOC, delta <= 4)
model.avg(bestSOC, subset = cumsum(weight) <= .95)
summary(get.models(bestSOC, 1)[[1]])

BestmodelSOC <- (get.models(bestSOC, 1)[[1]])
AICc(BestmodelSOC) # [1] 115.765 (Menor que el del full model, OK)
# Simplified summary
summ(BestmodelSOC, vifs=TRUE) # E2 and O2

#VIFS:
check_collinearity(BestmodelSOC) # = 1

#POST-HOCS:
emmeans (BestmodelSOC, pairwise ~ EarlyRearing) # E1-E2
emmeans (BestmodelSOC, pairwise ~ Origin) # O1 - O2
```
## RULE breaking 
            and EAE
```{r}
library(readxl)
DataRB <- read_excel("~/Desktop/Manuscrito/Data/Datafinal/nuevejul/databaseRB.xlsx")
```
```{r}
# Converting Rule Breaking to z-scores
z.RuleBreaking <- as.vector(scale(DataRB$RuleBreaking))

# Full model
fullRB <- glm(z.RuleBreaking ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = DataRB, na.action = "na.fail")
AICc(fullRB) # [1] 131.4474

# Best model
bestRB <- dredge(fullRB,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestRB, delta <= 4)
model.avg(bestRB, subset = cumsum(weight) <= .95)
summary(get.models(bestRB, 1)[[1]])

BestmodelRB <- (get.models(bestRB, 1)[[1]])
AICc(BestmodelRB) # [1] 111.3973 (Menor que el del full model, OK)
# Simplified summary
summ(BestmodelRB, vifs=TRUE) 

#VIFS:
check_collinearity(BestmodelRB) # <= 1.45 

#POST-HOCS:
emmeans (BestmodelRB, pairwise ~ MotherSeparation)
emmeans (BestmodelRB, pairwise ~ EarlyRearing) # E1-E2
```

## Aggressive behaviour 
            and EAE
```{r}
# Converting Aggressive behaviour to z-scores
z.AggressionRelated <- as.vector(scale(Data1$AggressionRelated))

# GLM 
# full model
fullAGGR <- glm(z.AggressionRelated ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullAGGR) # [1] 136.4169

# Best model
bestAGGR <- dredge(fullAGGR,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestAGGR, delta <= 4)
model.avg(bestAGGR, subset = cumsum(weight) <= .95)
summary(get.models(bestAGGR, 1)[[1]])

BestmodelAGGR <- (get.models(bestAGGR, 1)[[1]])
AICc(BestmodelAGGR) #[1] 110.064 (Menor que el del full model, OK)
# Simplified summary
summ(BestmodelAGGR) #

#VIFS:
# NO NEED

#POST-HOCS:
emmeans (BestmodelAGGR, pairwise ~ Origin) # O1 - O2
```
## Affiliative behaviour
                  and EAE
```{r}
library(readxl)
DataAFF <- read_excel("~/Desktop/Manuscrito/Data/Datafinal/nuevejul/databaseaff.xlsx")

# Converting Affiliative related behaviour to z-scores
z.Affiliative <- as.vector(scale(DataAFF$AffiliativeRelated))

# Full model
fullAFF <- glm(z.Affiliative ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = DataAFF, na.action = "na.fail")
AICc(fullAFF) # [1] 82.5701

# Best model
bestAFF <- dredge(fullAFF,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestAFF, delta <= 4)
model.avg(bestAFF, subset = cumsum(weight) <= .95)
summary(get.models(bestAFF, 1)[[1]])

BestmodelAFF <- (get.models(bestAFF, 1)[[1]])
AICc(BestmodelAFF) # [1] 63.17018 (Menor que el del full model, OK)
# Simplified summary
summ(BestmodelAFF, vifs=TRUE) # B4, E2, H2 and O2

#VIFS:
check_collinearity(BestmodelAFF)

#POST-HOCS:
emmeans (BestmodelAFF, pairwise ~ Abuse) 
emmeans (BestmodelAFF, pairwise ~ EarlyRearing) 
```

## Analytical
            and EAE
```{r}
# Converting Analytical to z-scores
z.Analytical <- as.vector(scale(Data1$Analytical))

# GLM 
# full model
fullANA <- glm(z.Analytical ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullANA) # 110.2169

# best fit model
bestANA <- dredge(fullANA,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestANA, delta <= 4)
model.avg(bestANA, subset = cumsum(weight) <= .95) 
summary(get.models(bestANA, 1)[[1]])

BestmodelANA <- (get.models(bestANA, 1)[[1]])
AICc(BestmodelANA) # [1] 107.4089
summ(BestmodelANA)

#VIFS:
check_collinearity(BestmodelANA)

#POST-HOCS:
emmeans (BestmodelANA, pairwise ~ Abuse) 
emmeans (BestmodelANA, pairwise ~ AgeArrival) 
```
## Cautiousness 
            and EAE
```{r}
# Converting Analytical to z-scores
z.Cautiousness <- as.vector(scale(Data1$Cautiousness))

# GLM 
# full model
fullCAU <- glm(z.Cautiousness ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullCAU) # [1] 138.7553

# best fit model
bestCAU <- dredge(fullCAU,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestCAU, delta <= 4)
model.avg(bestCAU, subset = cumsum(weight) <= .95) 
summary(get.models(bestCAU, 1)[[1]])

BestmodelCAU <- (get.models(bestCAU, 1)[[1]])
AICc(BestmodelCAU) # 120.9511
summ(BestmodelCAU)

#VIFS:
check_collinearity(BestmodelCAU)

#POST-HOCS:
emmeans (BestmodelCAU, pairwise ~ Origin) 
emmeans (BestmodelCAU, pairwise ~ AgeArrival) 
```
## Rule Consciousness
                and EAE
```{r}
# Converting Rule Consciousness to z-scores
z.RuleConsciousness <- as.vector(scale(Data1$RuleConsciousness))

# GLM 
# full model
fullRC <- glm(z.RuleConsciousness ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullRC) # [1] 140.0775

# best fit model
bestRC <- dredge(fullRC,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestRC, delta <= 4)
model.avg(bestRC, subset = cumsum(weight) <= .95) 
summary(get.models(bestRC, 1)[[1]])

BestmodelRC <- (get.models(bestRC, 1)[[1]])
AICc(BestmodelRC) # [1] 119.6669
summ(BestmodelRC)

#VIFS:
check_collinearity(BestmodelRC)

#POST-HOCS:
emmeans (BestmodelRC, pairwise ~ EarlyRearing) 
```
## Extraversion
            and EAE
```{r}
# Converting Extraversion to z-scores
z.Extraversion <- as.vector(scale(Data1$Extraversion))

# GLM 
# full model
fullEX <- glm(z.Extraversion ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullEX) # [1] 140.7581

# best fit model
bestEX <- dredge(fullEX,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestEX, delta <= 4)
model.avg(bestEX, subset = cumsum(weight) <= .95) 
summary(get.models(bestEX, 1)[[1]])

BestmodelEX <- (get.models(bestEX, 1)[[1]])
AICc(BestmodelEX) # [1] 119.2131
summ(BestmodelEX)

#POST-HOCS:
emmeans (BestmodelEX, pairwise ~ Origin) 
```

## Stab-Q-Chimp subtypes
             with EAE and 16Pf
# Physical aggression
```{r}
# full model
fullPHYtot <- glm(z.PhysicalAggression ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse + z.Analytical + z.Cautiousness + z.Extraversion + z.RuleConsciousness, data = Data1, na.action = "na.fail")
AICc(fullPHYtot) # [1] 120.3758

# best fit model
bestPHYtot <- dredge(fullPHYtot,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestPHYtot, delta <= 4)
model.avg(bestPHYtot, subset = cumsum(weight) <= .95) 
BestmodelPHYtot <- (get.models(bestPHYtot, 1)[[1]])
AICc(BestmodelPHYtot) # [1] 97.21521
summ(BestmodelPHYtot)
#VIFS:
check_collinearity(BestmodelPHYtot)

#POSTHOC
emmeans (BestmodelPHYtot, pairwise ~ Abuse) 
emmeans (BestmodelPHYtot, pairwise ~ AgeArrival)
```
# Social aggression
```{r}
# full model
fullSOCtot <- glm(z.SocialAggression ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse + z.Analytical + z.Cautiousness + z.Extraversion + z.RuleConsciousness, data = Data1, na.action = "na.fail")
AICc(fullSOCtot) # [1] 137.115

# best fit model
bestSOCtot <- dredge(fullSOCtot,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestSOCtot, delta <= 4)
model.avg(bestSOCtot, subset = cumsum(weight) <= .95) 
BestmodelSOCtot <- (get.models(bestSOCtot, 1)[[1]])
summary(get.models(bestSOCtot, 1)[[1]])
AICc(BestmodelSOCtot) # [1] 107.4076
summ(BestmodelSOCtot)
# VIF
check_collinearity(BestmodelSOCtot)

#POSTHOC
emmeans (BestmodelSOCtot, pairwise ~ Origin)
```
# Rule breaking
```{r}
#standarize 16PF in the DataRB
z.AnalyticalRB <- as.vector(scale(DataRB$Analytical))
z.CautiousnessRB <- as.vector(scale(DataRB$Cautiousness))
z.RuleConsciousnessRB <- as.vector(scale(DataRB$RuleConsciousness))
z.ExtraversionRB <- as.vector(scale(DataRB$Extraversion))

# full model
fullRBtot <- glm(z.RuleBreaking ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse + z.AnalyticalRB + z.CautiousnessRB + z.ExtraversionRB + z.RuleConsciousnessRB, data = DataRB, na.action = "na.fail")
AICc(fullRBtot) # [1] 149.5476

# best fit model
bestRBtot <- dredge(fullRBtot,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestRBtot, delta <= 4)
model.avg(bestRBtot, subset = cumsum(weight) <= .95) 
BestmodelRBtot <- (get.models(bestRBtot, 1)[[1]])
AICc(BestmodelRBtot) # [1] 111.3973
summ(BestmodelRBtot)
# VIF
check_collinearity(BestmodelRBtot)
#POSTHOC
emmeans (BestmodelRBtot, pairwise ~ EarlyRearing)
```

#########################################

# Plots - Visualization Data: Influence of EAE on STAB-Q-Chimp subtypes

# Prueba con ggplot2
```{r}
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("ggpubr")) install.packages("ggpubr")
library(ggplot2)
library(ggpubr)

library(multcomp)
library(ggsignif)
```
Bestmodel1
```{r}
Data1$Abuse <- as.factor(Data1$Abuse)
modelo <- glm(z.PhysicalAggression ~ Abuse, data = Data1)
resumen_posthoc <- summary(glht(modelo, linfct = mcp(Abuse = "Tukey")))

p_values <- resumen_posthoc$test$pvalues
comparisons <- list(
  c("None", "Low"),
  c("None", "Medium"),
  c("None", "High"),
  c("Low", "Medium"),
  c("Low", "High"),
  c("Medium", "High")
)

annotations <- ifelse(p_values < 0.001, "***",
                      ifelse(p_values < 0.01, "**",
                             ifelse(p_values < 0.05, "*", "ns")))
# Crear el gráfico con ggplot2
ggplot(Data1, aes(x = Abuse, y = z.PhysicalAggression, fill = Abuse)) +
  geom_boxplot(outlier.shape = 25, outlier.fill = "black", outlier.color = "black") +
  scale_fill_manual(values = colors1) +
  labs(title = "Best model 1 - Abuse",
       x = "Degree of Abuse",
       y = "Physical aggression") +
  scale_x_discrete(labels = c("None", "Low", "Medium", "High")) +
  theme_minimal() +
   geom_signif(comparisons = comparisons,
              annotations = annotations,  # Aquí se agregan los asteriscos de significancia
              map_signif_level = TRUE, 
              textsize = 6)

# Método gLM
ggplot(Data1, aes(x = Abuse, y = z.PhysicalAggression, fill = Abuse)) +
  geom_boxplot(outlier.shape = 25, outlier.fill = "black", outlier.color = "black") +
  scale_fill_manual(values = colors1) +
  labs(title = "Best model 1 - Abuse",
       x = "Degree of Abuse",
       y = "Physical aggression") +
  scale_x_discrete(labels = c("None", "Low", "Medium", "High")) +
  theme_minimal() +
  geom_signif(comparisons = comparisons, 
              map_signif_level = TRUE, 
              textsize = 6)


# Warning: [1m[22mComputation failed in `stat_compare_means()`
```


## BEST MODEL 1 - PHYSICAL AGGRESSION ##
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")

# Ventana gráfica a superponer los dos gráficos [ si hay solo un grafico: par(new = TRUE)]
par(mfrow = c(1,2)) # 1 fila y 2 columnas

# Best model 1: PHY - Abuse
boxplot(z.PhysicalAggression ~ Abuse, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Degree of Abuse",
        ylab = "Physical aggression",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("None", "Low", "Medium", "High"))

# Best model 1: PHY - AgeArrival
boxplot(z.PhysicalAggression ~ AgeArrival, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Age at arrival to the centre",
        ylab = "Physical aggression",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("0-2", "3-10", ">10", "captivity born"))

# Juntarlos en una sola imagen
par(mfrow = c(1, 1))
```
## BEST MODEL 2 - SOCIAL AGGRESSION ##
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(1,2))

# Best model 2: SOC - Early Rearing
boxplot(z.SocialAggression ~ EarlyRearing, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Early rearing conditions",
        ylab = "Social aggression",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Mother-reared", "Hand-reared"))
p + geom_signif(comparisons = list(c("Mother-reared", "Hand-reared")),
                map_signif_level = TRUE)
# Best model 2: SOC - Origin
boxplot(z.SocialAggression ~ Origin, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Origin",
        ylab = "Social aggression",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Captivity born", "Wildborn"))
par(mfrow = c(1, 1))
```
## BEST MODEL 3 - RULE BREAKING ##
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(new = TRUE)

# Best model 3: RB - Early Rearing
boxplot(z.RuleBreaking ~ EarlyRearing, data=DataRB, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Early rearing conditions",
        ylab = "Rule breaking",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Mother-reared", "Hand-reared")) #No hay unstructured, meter en la discusión (?)
```

# Plots - Visualization Data: Influence of EAE on Behavioural rates
## BEST MODEL 4 - AGGRESSIVE BEHAVIOUR ##
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(new = TRUE)

# Best model 4: AGGR - Origin
boxplot(z.AggressionRelated ~ Origin, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Origin",
        ylab = "Rate of aggressive behaviour",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Captivity born", "Wildborn"))
```
## BEST MODEL 5 - AFFILIATIVE BEHAVIOUR ##
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(1,2))

# Best model 5: AFF - Abuse
boxplot(z.Affiliative ~ Abuse, data = DataAFF, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Degree of abuse",
        ylab = "Rate of affiliative behaviour",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("None", "Low", "Medium", "High"))

# Best model 5: AFF - Early rearing
boxplot(z.Affiliative ~ EarlyRearing, data = DataAFF, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Early rearing conditions",
        ylab = "Rate of affiliative behaviour",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Mother-reared", "Hand-reared"))

par(mfrow = c(1, 1))
```

# Plots - Visualization Data: Influence of EAE on personality 16PF dimensions
## BESTMODEL 6 - ANALYTICAL
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(1,2))

# Best model 6: ANA - Abuse
boxplot(z.Analytical ~ Abuse, data = Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Degree of abuse",
        ylab = "Analytical levels",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("None", "Low", "Medium", "High"))

# Best model 6: ANA - Age at arrival
boxplot(z.Analytical ~ AgeArrival, data = Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Age at arrival to the centre",
        ylab = "Analytical levels",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("0-2", "3-10", ">10", "captivity born"))

par(mfrow = c(1, 1))
```
CAUTIOUSNESS NO TIENE VARIABLE PREDICTORA SIGNIFICATIVA
## BESTMODEL 8 - RULE-CONSCIOUSNESS
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(new = TRUE)

# Best model 4: CONSC - Early rearing
boxplot(z.RuleConsciousness ~ EarlyRearing, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Early rearing conditions",
        ylab = "Rule Consciousness levels",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Mother-reared", "Hand-reared"))


```
## BESTMODEL 9 - EXTRAVERSION
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(new = TRUE)

# Best model 4: EXTRAV - Origin
boxplot(z.AggressionRelated ~ Origin, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Origin",
        ylab = "Extraversion levels",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Captivity born", "Wildborn"))
```

# Plots - Visualization Data: Influence of EAE + 16PF on STAB-Q-Chimp subtypes
## BESTMODEL 10 - PHYSICAL AGGRESSION
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(2,2))

# Best model 10: PHY - Abuse
boxplot(z.PhysicalAggression ~ Abuse, data = Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Degree of abuse",
        ylab = "Physical aggression",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("None", "Low", "Medium", "High"))

# Best model 10: PHY - Age at arrival
boxplot(z.PhysicalAggression ~ AgeArrival, data = Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Age at arrival to the centre",
        ylab = "Physical aggression",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("0-2", "3-10", ">10", "captivity born"))

# Bestmodel 10: PHY - Cautiousness
plot(Data1$Cautiousness, z.PhysicalAggression,
     col = "grey",  
     pch = 19,  
     xlab = "Cautiousness",  
     ylab = "Physical Aggression") 
lines(lowess(Data1$Cautiousness, z.PhysicalAggression), col = "#EE3B3B", lwd = 2, lty = 2)
abline(lm(z.PhysicalAggression ~ Cautiousness, data = Data1), col = "black", lwd = 1)

# Bestmodel 10: PHY - Rule-Consciousness
plot(Data1$RuleConsciousness, z.PhysicalAggression,
     col = "grey", 
     pch = 19,  
     xlab = "Rule Consciousness",  
     ylab = "Physical Aggression")  
lines(lowess(Data1$RuleConsciousness, z.PhysicalAggression), col = "#EE3B3B", lwd = 2, lty = 2)
abline(lm(z.PhysicalAggression ~ RuleConsciousness, data = Data1), col = "black", lwd = 1) 

par(mfrow = c(1, 1))
```
## BestMODEL 11 - SOCIAL AGGRESSION
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(2,2))

# Best model 11: SOC - Origin
boxplot(z.SocialAggression ~ Origin, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Origin",
        ylab = "Social aggression",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Captivity born", "Wildborn"))

# Best model 11: SOC - Analytical
plot(Data1$Analytical, z.SocialAggression,
     col = "grey",  
     pch = 19,  
     xlab = "Analytical",  
     ylab = "Social Aggression") 
lines(lowess(Data1$Analytical, z.SocialAggression), col = "#EE3B3B", lwd = 2, lty = 2)
abline(lm(z.SocialAggression ~ Analytical, data = Data1), col = "black", lwd = 1) 

# Best model 11: SOC - Cautiousness
plot(Data1$Cautiousness, z.SocialAggression,
     col = "grey",  
     pch = 19,  
     xlab = "Cautiousness",  
     ylab = "Social Aggression") 
lines(lowess(Data1$Cautiousness, z.SocialAggression), col = "#EE3B3B", lwd = 2, lty = 2)
abline(lm(z.SocialAggression ~ Cautiousness, data = Data1), col = "black", lwd = 1)

# Best model 11: SOC - Rule-Consciousness
plot(Data1$RuleConsciousness, z.SocialAggression,
     col = "grey", 
     pch = 19,  
     xlab = "Rule Consciousness",  
     ylab = "Social Aggression")  
lines(lowess(Data1$RuleConsciousness, z.SocialAggression), col = "#EE3B3B", lwd = 2, lty = 2)
abline(lm(z.SocialAggression ~ RuleConsciousness, data = Data1), col = "black", lwd = 1) 

par(mfrow = c(1, 1))
```
## BESTMODEL 12 - RULE BREAKING (Es exactamente lo mismo que el anterior Best model 2 pero solo con early rearing)
```{r}
 colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(new = TRUE)

boxplot(z.RuleBreaking ~ EarlyRearing, data=DataRB, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Early rearing conditions",
        ylab = "Rule breaking",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Mother-reared", "Hand-reared"))
```

