---
title: "GlmsRedefined"
output: html_document
date: "2025-10-11"
---


```{r setup, include=FALSE}
# MuMIn: For model selection based on Akaike Criterion (AIC)
if (!require("MuMIn")) install.packages("MuMIn")
# car for collinearity
if (!require("car")) install.packages("car")
# formatting nas
if (!require("dplyr")) install.packages("dplyr")
# colinearity with checkcolinearity
if (!require("performance")) install.packages("performance")
# summary with j.tools
if (!require("jtools")) install.packages("jtools")
# for post-hocs
if (!require("emmeans")) install.packages("emmeans")

# Package loading
library(MuMIn)
library(car)
library(dplyr)
library(performance) # checkcollinearity
library(jtools) #summ
library(emmeans) #post-hocs
```

## Loading data
```{r}
library(readxl)
Data1 <- read_excel("yourroute")
View(Data1)
```

## PHYSICAL Aggression 
            and EAE
```{r}
# Converting Physical Aggression to z-scores
z.PhysicalAggression <- as.vector(scale(Data1$PhysicalAggression))

# Full model
fullPHY <- glm(z.PhysicalAggression ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullPHY) # [1] 136.959
check_collinearity(fullPHY) # >10

# Best model
bestPHY <- dredge(fullPHY,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestPHY, delta <= 4)
model.avg(bestPHY, subset = cumsum(weight) <= .95)
summary(get.models(bestPHY, 1)[[1]]) # Early Rearing E2 is significant only.

BestmodelPHY <- (get.models(bestPHY, 1)[[1]])
AICc(BestmodelPHY) # [1] 116.986 OK

# Simplified summary
summ(BestmodelPHY)

#VIFS:
check_collinearity(BestmodelPHY) # = 1.02 OK

#POST-HOCS:
emmeans (BestmodelPHY, pairwise ~ EarlyRearing)
```

## DISRUPTIVE Behavior
            and EAE
```{r}
# Converting Disruptive Behavior to z-scores
z.DisruptiveBehavior <- as.vector(scale(Data1$DisruptiveBehavior))

# Full model
fulldis <- glm(z.DisruptiveBehavior ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fulldis) # [1] 131.6923
check_collinearity(fulldis) # >10

# Best model
bestdis <- dredge(fulldis,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestdis, delta <= 4)
model.avg(bestdis, subset = cumsum(weight) <= .95)
summary(get.models(bestdis, 1)[[1]]) # Intercept and Early Rearing E2 significant

Bestmodeldis <- (get.models(bestdis, 1)[[1]])
AICc(Bestmodeldis) # [1] 113.949 OK

#VIFS: Not needed

#POST-HOCS:
emmeans (Bestmodeldis, pairwise ~ EarlyRearing)
```

## Aggressive behaviour 
            and EAE
```{r}
# Converting Aggressive behaviour to z-scores
z.AggressionRelated <- as.vector(scale(Data1$Aggressionrelated))
 
# Full model
fullAGGR <- glm(z.AggressionRelated ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullAGGR) # [1] 139.972
check_collinearity(fullAGGR) # >10

# Best model
bestAGGR <- dredge(fullAGGR,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestAGGR, delta <= 4)
model.avg(bestAGGR, subset = cumsum(weight) <= .95)
summary(get.models(bestAGGR, 1)[[1]]) # Origin O2 significant

BestmodelAGGR <- (get.models(bestAGGR, 1)[[1]])
AICc(BestmodelAGGR) #[1] 114.108 OK

# Simplified summary
summ(BestmodelAGGR)

#VIFS: Not needed

#POST-HOCS:
emmeans (BestmodelAGGR, pairwise ~ Origin) 
```

## Affiliative behaviour
                   and EAE
```{r}
library(readxl)
DataAFF <- read_excel("routetodatabaseFORaffiliativebehavior")

# Converting Affiliative related behaviour to z-scores
z.Affiliative <- as.vector(scale(DataAFF$AffiliativeRelated))

# Full model
fullAFF <- glm(z.Affiliative ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = DataAFF, na.action = "na.fail")
AICc(fullAFF) # [1] 82.5701

# Best model
bestAFF <- dredge(fullAFF,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestAFF, delta <= 4)
model.avg(bestAFF, subset = cumsum(weight) <= .95)
summary(get.models(bestAFF, 1)[[1]]) # Intercept, Abuse B3, and Early Rearing E2

BestmodelAFF <- (get.models(bestAFF, 1)[[1]])
AICc(BestmodelAFF) # [1] 63.17018 OK

# Simplified summary
summ(BestmodelAFF, vifs=TRUE)

#VIFS:
check_collinearity(BestmodelAFF) OK

#POST-HOCS:
emmeans (BestmodelAFF, pairwise ~ Abuse) 
emmeans (BestmodelAFF, pairwise ~ Origin) 
```

## Analytical
            and EAE
```{r}
# Converting Analytical to z-scores
z.Analytical <- as.vector(scale(Data1$Analytical))

# Full model
fullANA <- glm(z.Analytical ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullANA) # [1] 121.4673
check_collinearity(fullANA)

# Best model
bestANA <- dredge(fullANA,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestANA, delta <= 4)
model.avg(bestANA, subset = cumsum(weight) <= .95) 
summary(get.models(bestANA, 1)[[1]]) #Mother Separation M2 and M3

BestmodelANA <- (get.models(bestANA, 1)[[1]])
AICc(BestmodelANA) # [1] 107.4089 OK

# Simplified summary
summ(BestmodelANA)

#VIFS:
check_collinearity(BestmodelANA)

#POST-HOCS:
emmeans (BestmodelANA, pairwise ~ MotherSeparation) 
emmeans (BestmodelANA, pairwise ~ PastExperiences) 
```

## Extraversion
            and EAE
```{r}
# Converting Extraversion to z-scores
z.Extraversion <- as.vector(scale(Data1$Extraversion))

# GLM 
# Full model
fullEXTR <- glm(z.Extraversion ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullEXTR) # 146.3017
check_collinearity(fullEXTR)

# Best model
bestEXTR <- dredge(fullEXTR,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestEXTR, delta <= 4)
model.avg(bestEXTR, subset = cumsum(weight) <= .95) 
summary(get.models(bestEXTR, 1)[[1]]) # none

BestmodelEXTR <- (get.models(bestEXTR, 1)[[1]])
AICc(BestmodelEXTR) # [1] 122.4864 OK

# Simplified summary
summ(BestmodelEXTR)
```

## Withdrawal
            and EAE
```{r}
# Converting Withdrawal to z-scores
z.Withdrawal <- as.vector(scale(Data1$Withdrawal))

# Full model
fullWITHD <- glm(z.Withdrawal ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullWITHD) # 148.9611

# Best model
bestWITHD <- dredge(fullWITHD,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestWITHD, delta <= 4)
model.avg(bestWITHD, subset = cumsum(weight) <= .95) 
summary(get.models(bestWITHD, 1)[[1]]) # None

BestmodelWITHD <- (get.models(bestWITHD, 1)[[1]])
AICc(BestmodelWITHD) # [1] 122.4864

# Simplified summary
summ(BestmodelWITHD)
```

## Affiliation
            and EAE
```{r}
# Converting Withdrawal to z-scores
z.Affiliation <- as.vector(scale(Data1$Affiliation))

# Full model
fullAFFIL <- glm(z.Affiliation ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse, data = Data1, na.action = "na.fail")
AICc(fullAFFIL) # 138.3486
check_collinearity(fullAFFIL)

# Best model
bestAFFIL <- dredge(fullAFFIL,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestAFFIL, delta <= 4)
model.avg(bestAFFIL, subset = cumsum(weight) <= .95) 
summary(get.models(bestAFFIL, 1)[[1]]) 

BestmodelAFFIL <- (get.models(bestAFFIL, 1)[[1]])
AICc(BestmodelAFFIL) # [1] 121.7247 OK

# Simplified summary
summ(BestmodelAFFIL)
```
# Combination:
    16PF + EAE on predicting STAB-Q-Chimp factors
# Physical aggression

```{r}
# Full model
fullPHYtot <- glm(z.PhysicalAggression ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse + z.Analytical + z.Extraversion + z.Withdrawal + z.Affiliation, data = Data1, na.action = "na.fail")
AICc(fullPHYtot) # [1] 138.9542
check_collinearity(fullPHYtot)

# Best model
bestPHYtot <- dredge(fullPHYtot,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestPHYtot, delta <= 4)
model.avg(bestPHYtot, subset = cumsum(weight) <= .95) 
BestmodelPHYtot <- (get.models(bestPHYtot, 1)[[1]])
AICc(BestmodelPHYtot) # 110.5316
summary(get.models(bestPHYtot, 1)[[1]])

# Simplified summary
summ(BestmodelPHYtot)

#VIFS:
check_collinearity(BestmodelPHYtot) # = 1.02 

#POSTHOC
emmeans (BestmodelPHYtot, pairwise ~ EarlyRearing) 
```

# Disruptive behavior

```{r}
# Full model
fullDISRtot <- glm(z.DisruptiveBehavior ~ Origin + EarlyRearing + AgeArrival + PastExperiences + MotherSeparation + Abuse + z.Analytical + z.Extraversion + z.Withdrawal + z.Affiliation, data = Data1, na.action = "na.fail")
AICc(fullDISRtot) # [1] 127.8653
check_collinearity(fullDISRtot)

# Best model
bestDISRtot <- dredge(fullDISRtot,  rank = "AICc", trace=4, fixed = NULL, m.lim = NULL)
subset(bestDISRtot, delta <= 4)
model.avg(bestDISRtot, subset = cumsum(weight) <= .95) 
BestmodelDISRtot <- (get.models(bestDISRtot, 1)[[1]])
AICc(BestmodelDISRtot) # 103.2769 OK
summary(get.models(bestDISRtot, 1)[[1]])

# Simplified summary
summ(BestmodelDISRtot)

#VIFS:
check_collinearity(BestmodelDISRtot) # = 1.02 

#POSTHOC
emmeans (BestmodelDISRtot, pairwise ~ EarlyRearing) 
```

# Plots - Visualization Data: Influence of EAE on STAB-Q-Chimp subtypes

# Prueba con ggplot2
```{r}
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("ggpubr")) install.packages("ggpubr")
library(ggplot2)
library(ggpubr)

library(multcomp)
library(ggsignif)
```

# Best model 1 - Physical Aggression
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(1,2))

# Best model 1: PHY - Early Rearing
dev.new(width = 10, height = 10)
boxplot(z.PhysicalAggression ~ EarlyRearing, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Early Rearing Conditions",
        ylab = "Physical aggression",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Mother-reared", "Hand-reared"))
```

# Best model 2 - Disruptive Behavior
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(1,2))

# Best model 3: DB - Early Rearing
dev.new(width = 10, height = 6)
boxplot(z.DisruptiveBehavior ~ Origin, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Early Rearing Conditions",
        ylab = "Disruptive Behavior",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Mother-reared", "Hand-reared"))
```

# Best model 3 - Aggressive Behavior
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(1,2))

# Best model 3: AB - Origin
boxplot(z.AggressionRelated ~ Origin, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Origin",
        ylab = "Aggressive Behavior",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Captivity born", "Wildborn"))
```

# Best model 4 - Affiliative Behavior
```{r}
# Introduction same as before.
# Best model 4: AFF - Abuse
dev.new(width = 10, height = 6)
boxplot(z.Affiliative ~ Abuse, data=DataAFF, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Degree of Abuse",
        ylab = "Affiliative Behavior",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("None", "Low", "Medium", "High"))

# Best model 4: AFF - Origin
boxplot(z.Affiliative ~ Origin, data=DataAFF, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Origin",
        ylab = "Affiliative Behavior",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Captivity born", "Wildborn"))

par(mfrow = c(1, 1))
```

# Best model 5 - Analytical
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(2,2))

# Best model 5: ANA - Mother Separation
dev.new(width = 10, height = 6)
boxplot(z.Analytical ~ MotherSeparation, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Mother Separation before 5 years old",
        ylab = "Analytical-Concrete",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Yes", "No", "Unknown"))

# Best model 5: ANA - Past Experiences
dev.new(width = 15, height = 6)
par(mar = c(12, 5, 4, 2))
boxplot(z.Analytical ~ PastExperiences, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Past Experiences",
        ylab = "Analytical-Concrete",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Bushmeat", "Pet and Performer", "Zoo-born", "Sanctuary-born"))

par(mfrow = c(1, 1))
```

# Best model 9 - PHYSICAL AGGRESSION
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(2,2))

# Best model 9 - PHY and Early Rearing 
boxplot(z.PhysicalAggression ~ EarlyRearing, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Early Rearing Conditions",
        ylab = "Physical Aggression",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Mother-reared", "Hand-reared"))

# Best model 5 - PHY and Affiliation
plot(Data1$Affiliation, z.PhysicalAggression,
     col = "grey",  
     pch = 19,  
     xlab = "Affiliation",  
     ylab = "Physical Aggression") 
lines(lowess(Data1$Affiliation, z.PhysicalAggression), col = "#EE3B3B", lwd = 2, lty = 2)
abline(lm(z.PhysicalAggression ~ Affiliation, data = Data1), col = "black", lwd = 1) 

# Combine both plots
par(mfrow = c(1, 1))
```

# Best model 10 - DISTRUPTIVE BEHAVIOR 
```{r}
colors1 = c("#C1CDCD", "#EE3B3B", "#8EE5EE", "#BCEE68")
plot.new() #Crear gráfico en planco
set.seed(1)
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "#FDFAFA")
par(mfrow = c(2,2))

# Best model 10 - DB and Early Rearing 
boxplot(z.DisruptiveBehavior ~ EarlyRearing, data=Data1, horizontal = FALSE,
        lwd = 1,
        col = colors1,
        xlab = "Early Rearing Conditions",
        ylab = "Disruptive Behavior",
        border = "black",
        outpch = 25,
        outbg = "black",
        whiskcol = "black",
        whisklty = 1,
        lty = 1,
        names = c("Mother-reared", "Hand-reared"))

# Best model 5 - PHY and Affiliation
plot(Data1$Affiliation, z.DisruptiveBehavior,
     col = "grey",  
     pch = 19,  
     xlab = "Affiliation",  
     ylab = "Disruptive Behavior") 
lines(lowess(Data1$Affiliation, z.DisruptiveBehavior), col = "#EE3B3B", lwd = 2, lty = 2)
abline(lm(z.DisruptiveBehavior ~ Affiliation, data = Data1), col = "black", lwd = 1) 

# Combine both plots
par(mfrow = c(1, 1))
```
